---
title: "Prediction Assignment Writeup"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, error=TRUE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = FALSE)
```


```{r Packages, results='hide',message=FALSE}
library(caret);library(rpart);library(randomForest)
```

Setting seed for reproducibility

```{r}
set.seed(45)
```

Loading the data
```{r Preprocessing}
training <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",na.strings=c("NA","#DIV/0!",""))
```

We are going to eliminate the columns that contains NA values
```{r}
training<-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]
```
Also, the first 5 columns are irrelevant to be considered as predictors so we remove them
```{r}
training   <-training[,-c(1:5)]
testing <-testing[,-c(1:5)]

#Final Dimensions of the data
dim(training)
dim(testing)
```

```{r}
#COnverting Classe to factor
training$classe <-factor(training$classe)
```


Cross-validation
```{r cross}
#In order to get out-of-sample errors, split the training data in training (75%) and testing (25%) data) subsets:
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)    
datatrain <- training[inTrain, ]
datatest <- training[-inTrain, ]  
dim(datatrain)
dim(datatest)
```


Machine Learning Algorithms

Here we are going to use 2 machine learning algo

1)Decision Trees
2)Random Forest



```{r model}
#DECISION TREE
#Fit model on datatrain data
fitDT <- rpart(classe ~ ., data= datatrain, method="class")

#Use model to predict class in validation set (datatest)
predictionDT <- predict(fitDT, datatest, type = "class")

#Estimate the errors of the prediction algorithm in the Decision Tree model
confusionMatrix(datatest$classe, predictionDT)

#RANDOM FOREST
#Fit model on datatrain data
fitRF <- randomForest(classe ~ ., data=datatrain, method="class")

#Use model to predict class in validation set (datatest)
predictionRF <- predict(fitRF, datatest, type = "class")

#Estimate the errors of the prediction algorithm in the Random Forest
confusionMatrix(datatest$classe, predictionRF)
```

TEST THE MODEL TO PREDICT 20 DIFFERENT TEST CASES
```{r quiz}
# Perform prediction
predictSubmission <- predict(fitRF, testing, type="class")
predictSubmission
```